{\fontfamily{ptm}\selectfont Nesterov Type Momentum Methods}

### 2024-08-02 Notebook LaTEX/Momentum Accelerations/main.tex
>w<, aaa

            \begin{definition}[Similar Triangle Form I]
            \label{def:ag_form_similar_tria_I}
                It updates iterates $(y_t, x_{t + 1}, z_{t + 1})$ using 
                \begin{align*}
                    z_{t + 1} &= y_t - L^{-1} \mathcal G_L(y_t) 
                    \\
                    x_{t + 1} &= z_{t + 1} + L\eta_t (z_{t + 1} - z_t)
                    \\
                    y_{t + 1} &= 
                    (1 + L\eta_{t + 1})^{-1}
                    (
                    x_{t + 1} + L\eta_{t + 1}z_{t + 1}
                    ). 
                \end{align*}
            \end{definition}
            \begin{observation}
                There are a lot of observations about this form to unpack. 
                Let's list the following observations
                \begin{enumerate}
                    \item Observe that $y_t = (1 + L\eta_t)^{-1}(x_t + L\eta_t z_t)$ is a convex combination between $x_t, z_t$. Hence $z_t, y_t, x_t$ are three collinear points. Additionally, we have $y_t - x_t = L\eta_t (z_t - y_t)$, hence $\Vert y_t - x_t\Vert/\Vert z_t - y_t\Vert = L\eta_t$. 
                    \item By the algorithm, we have $z_{t + 1} - y_t = - L^{-1} \nabla f(y_t)$, and $x_{t + 1} - x_t = - \tilde \eta_{t + 1} \nabla f(y_t)$, hence vector $z_{t + 1} - y_t$ parallels to $x_{t + 1} - x_t$. 
                    \item Finally, with update $x_{t +1} - z_{t + 1} = L\eta_t (z_{t +1} - z_t)$, we have three colinear points $(z_t, z_{t + 1}, x_t)$ with $\Vert z_{t + 1} - z_t\Vert/\Vert z_{t +1 } - z_t\Vert = L\eta_t$. 
                \end{enumerate}
                From the above results, we can conclude that triangle $(y_t, z_t, z_{t + 1})$ is similar to $(x_t, z_t, x_{t + 1})$ because they share colinear points $z_t, y_t, x_t$ and $z_t, z_{t + 1}, x_{t + 1}$, and their sides $z_{t + 1} - y_t$ parallels to $x_{t + 1} - x_t$. 
            \end{observation}
            \begin{remark}
                For more information about the specific choice of $\eta_t$ allows for an optimal convergence rates, see 
                \hyperref[sec:scenario_ii]{section \ref*{sec:scenario_ii}} for more information. 
            \end{remark}

@@

        \begin{lemma}[The Lyapunov function from estimating sequence and function]
            \;\\
            Assume that $f$ bounded below with $f_* = \min_x f(x)$ and there exists minimizer $x_*$. 
            With $\phi_k$ being an estimating functions and convex, $\{x_k\}$ an estimating sequence for $\phi_k$, then it satisfies
            \begin{align*}
                f(x_k) - f_*
                &\le 
                \lambda_k(\phi_0(x_*) - f_*). 
            \end{align*}
            for all $k\in \mathbb N\cup \{0\}$. 
        \end{lemma}
        \begin{proof}
            Directly by the property of estimating sequence $x_k$, we have for all $k, x$
            \begin{align*}
                \phi_k(x) - f(x) \ge \phi_k^* - f(x) \ge f(x_k) - f(x), 
            \end{align*}
            so 
            \begin{align*}
                f(x_k) - f(x) \le \lambda_k (\phi_0(x) - f(x))
                \\
                \implies 
                f(x_k) - f_* \le \lambda_k (\phi_0(x_*) - f(x_*)). 
            \end{align*}
            This establishes the convergence rate of $f(x_k)$ assuming that $\lambda_k$ is sequence that converges to zero. 
        \end{proof}
    \subsection{Construction of the estimating sequence and function}
        The construction of the Nesterov estimating sequence and function is where most creativities are involved. 
        For convex $f$ that is differentiable with $L$-Lipschitz gradient and strong convexity constant $\mu \ge 0$, Nesterov suggested in his book \cite{nesterov_lectures_2018} that there exists sequence $y_k$, $\gamma_k$, where
        \begin{align*}
            l_f(x, y_k) &= 
            f(y_k) + \langle \nabla f(y_k), x - y_k\rangle + 
            \frac{\mu}{2}\Vert x - y_k\Vert^2
            \\
            \phi_0 (x) &= 
            f(x_0) + \frac{\gamma_0}{2}\Vert x - v_0\Vert^2
            \\
            \phi_{k + 1}(x) &= 
            (1 - \alpha_k)\phi_k(x) + \alpha_k l_f(x; y_k)
            \\
            &= \phi_{k + 1}^2 + \frac{\gamma_k}{2}\Vert x - v_{k + 1}\Vert^2
        \end{align*}
        will verify as a Nesterov's estimating sequence and functions. 
        The derivation for the parameters: $\gamma_k, x_k, y_k$ for $k \in \mathbb N \cup \{0\}$, will produce the Nesterov acclerated gradient as presented in
        \hyperref[def:Nes2.2.7]{definition \ref*{def:Nes2.2.7}}
        if correctly conducted. 
        In this case, $\phi_k$ are simple quadratic centered $v_k$ with value $\phi_k$ with curvature $\gamma_k$. 
        \par
        However, the construction exhibits variations for different variants of the accelerated gradient algorithm. 
        A different construction of $\phi_k$ yield a different algorithm. 
        As shown in Nesterov's book \cite*[(2.2.63)]{nesterov_lectures_2018}, a projected gradient variant of the accelerated gradient method make use of the following estimating functions, sequence: 
        \begin{align*}
            \phi_0 (x) &= f(x_0) + \frac{\gamma_0}{2}\Vert x - x_0\Vert^2
            \\
            l_f(x; y_k) &= f(\mathcal T_ly_k) + 
            \frac{1}{2L}\Vert \mathcal G_L(y_k)\Vert^2 + 
            \langle \mathcal G_L(y_k), x - y_k\rangle + 
            \frac{\mu}{2}\Vert x - y_k\Vert^2
            \\
            \phi_{k + 1}(x) &= 
            (1 - \alpha_k)\phi_k + 
            \alpha_k l_f(x; y_k)
        \end{align*}

>>>>

            \begin{definition}[Similar Triangle Form I]
            \label{def:ag_form_similar_tria_I}
                It updates iterates $(y_t, x_{t + 1}, z_{t + 1})$ using 
                \begin{align*}
                    z_{t + 1} &= y_t - L^{-1} \mathcal G_L(y_t) 
                    \\
                    x_{t + 1} &= z_{t + 1} + L\eta_t (z_{t + 1} - z_t)
                    \\
                    y_{t + 1} &= 
                    (1 + L\eta_{t + 1})^{-1}
                    (x_{t + 1} + L\eta_{t + 1}z_{t + 1}). 
                \end{align*}
            \end{definition}
            \begin{observation}
                There are a lot of observations about this form to unpack. 
                Let's list the following observations
                \begin{enumerate}
                    \item Observe that $y_t = (1 + L\eta_t)^{-1}(x_t + L\eta_t z_t)$ is a convex combination between $x_t, z_t$. Hence $z_t, y_t, x_t$ are three collinear points. Additionally, we have $y_t - x_t = L\eta_t (z_t - y_t)$, hence $\Vert y_t - x_t\Vert/\Vert z_t - y_t\Vert = L\eta_t$. 
                    \item By the algorithm, we have $z_{t + 1} - y_t = - L^{-1} \mathcal G_L(y_t)$, and $x_{t + 1} - x_t = - \tilde \eta_{t + 1} \mathcal G_L(y_t)$, hence vector $z_{t + 1} - y_t$ parallels to $x_{t + 1} - x_t$. 
                    \item Finally, with update $x_{t +1} - z_{t + 1} = L\eta_t (z_{t +1} - z_t)$, we have three colinear points $(z_t, z_{t + 1}, x_t)$ with $\Vert z_{t + 1} - z_t\Vert/\Vert z_{t +1 } - z_t\Vert = L\eta_t$. 
                \end{enumerate}
                From the above results, we can conclude that triangle $(y_t, z_t, z_{t + 1})$ is similar to $(x_t, z_t, x_{t + 1})$ because they share colinear points $z_t, y_t, x_t$ and $z_t, z_{t + 1}, x_{t + 1}$, and their sides $z_{t + 1} - y_t$ parallels to $x_{t + 1} - x_t$. 
            \end{observation}
            \begin{remark}
                For more information about the specific choice of $\eta_t$ allows for an optimal convergence rates, see 
                \hyperref[sec:scenario_ii]{section \ref*{sec:scenario_ii}} for more information. 
            \end{remark}

@@

        \begin{lemma}[The Lyapunov function from estimating sequence and function]
            \;\\
            Assume that $f$ bounded below with $f_* = \min_x f(x)$ and there exists minimizer $x_*$. 
            With $\phi_k$ being an estimating functions and convex, $\{x_k\}$ an estimating sequence for $\phi_k$, then it satisfies
            \begin{align*}
                f(x_k) - f_*
                &\le 
                \lambda_k(\phi_0(x_*) - f_*). 
            \end{align*}
            for all $k\in \mathbb N\cup \{0\}$. 
        \end{lemma}
        \begin{proof}
            Directly by the property of estimating sequence $x_k$, we have for all $k, x$
            \begin{align*}
                \phi_k(x) - f(x) \ge \phi_k^* - f(x) \ge f(x_k) - f(x), 
            \end{align*}
            so 
            \begin{align*}
                f(x_k) - f(x) \le \lambda_k (\phi_0(x) - f(x))
                \\
                \implies 
                f(x_k) - f_* \le \lambda_k (\phi_0(x_*) - f(x_*)). 
            \end{align*}
            This establishes the convergence rate of $f(x_k)$ assuming that $\lambda_k$ is sequence that converges to zero. 
        \end{proof}
    \subsection{Construction of the estimating sequence and function}
        The construction of the Nesterov estimating sequence and function is where most creativities are involved. 
        For convex $f$ that is differentiable with $L$-Lipschitz gradient and strong convexity constant $\mu \ge 0$, Nesterov suggested in his book \cite{nesterov_lectures_2018} that there exists sequence $y_k$, $\gamma_k$, where
        \begin{align*}
            l_f(x, y_k) &= 
            f(y_k) + \langle \nabla f(y_k), x - y_k\rangle + 
            \frac{\mu}{2}\Vert x - y_k\Vert^2
            \\
            \phi_0 (x) &= 
            f(x_0) + \frac{\gamma_0}{2}\Vert x - v_0\Vert^2
            \\
            \phi_{k + 1}(x) &= 
            (1 - \alpha_k)\phi_k(x) + \alpha_k l_f(x; y_k)
            \\
            &= \phi_{k + 1}^2 + \frac{\gamma_k}{2}\Vert x - v_{k + 1}\Vert^2
        \end{align*}
        will verify as a Nesterov's estimating sequence and functions. 
        The derivation for the parameters: $\gamma_k, x_k, y_k$ for $k \in \mathbb N \cup \{0\}$, will produce the Nesterov acclerated gradient as presented in
        \hyperref[def:Nes2.2.7]{definition \ref*{def:Nes2.2.7}}
        if correctly conducted. 
        In this case, $\phi_k$ are simple quadratic centered $v_k$ with value $\phi_k$ with curvature $\gamma_k$. 
        \par
        However, the construction exhibits variations for different variants of the accelerated gradient algorithm. 
        A different construction of $\phi_k$ yield a different algorithm. 
        As shown in Nesterov's book \cite*[(2.2.63)]{nesterov_lectures_2018}, a projected gradient variant of the accelerated gradient method make use of the following estimating functions, sequence: 
        \begin{align*}
            \phi_0 (x) &= f(x_0) + \frac{\gamma_0}{2}\Vert x - x_0\Vert^2
            \\
            l_f(x; y_k) &= f(\mathcal T_ly_k) + 
            \frac{1}{2L}\Vert \mathcal G_L(y_k)\Vert^2 + 
            \langle \mathcal G_L(y_k), x - y_k\rangle + 
            \frac{\mu}{2}\Vert x - y_k\Vert^2
            \\
            \phi_{k + 1}(x) &= 
            (1 - \alpha_k)\phi_k + 
            \alpha_k l_f(x; y_k), 
        \end{align*}
        Here, $G_L(x) = L(x - \mathcal T_L(x))$ and $T_L(x) = \Pi_Q(x - L^{-1}\nabla f(x))$, where $Q$ is a convex set. 
        It's equivalent to applying proximal gradient method with $h = f + g$ where $g = \delta_Q$.
